\subsection*{MathML in HTML5 - Implementation Note\\Editor's Draft, \today}

\textbf{Editors:} \\
Frédéric Wang

\subsection*{Abstract}

This document is a detailed technical document for a core subset of presentation
MathML \cite{MathML3} in order to render with high quality the majority of
mathematical formulas used in practice.
This document should be understandable by any implementer with
sufficient knowledge of web rendering engines and text layout.
Contrary to the official MathML specification, a significant effort is made to
be as accurate as possible on the visual rendering of mathematical formulas
using additional rules from the \TeX Book's Appendix G \cite{TeXBook} and from
the Open Font Format version 3 \cite{OpenFontFormat3}.
The priority is to be compatible with existing technologies of web
rendering engines \cite{HTML5} by relying as much as possible on CSS, text \&
table layout and box models. As a consequence, parts of presentation MathML
that do not fit well in this framework or are rarely used in practice
have been ommited ; details on these and suggestions for standardization bodies
are provided in the appendix.

\subsection*{Status of This document}

This section describes the status of this document at the time of its
publication. Other documents may supersede this document.

This is a draft document and may be updated, replaced or obsoleted by other
documents at any time. It is inappropriate to cite this document as other than
work in progress.

\subsection*{Table of Contents}

\renewcommand\contentsname{}
\tableofcontents
